---
title: "variation"
author: "Jovani Ntabgoba"
date: "May 3, 2016"
output: html_document
---
To understand anything you must decipher patterns of variation.
The best way to study variation is to collect data, particularly rectangular data: data that is made up of variables, observations, and values.
        * A variable is a quantity, quality, or property that you can measure.
        * A value is the state of a variable when you measure it. The value of a variable may change from measurement to measurement.
        *An observation is a set of measurements you make under similar conditions (usually all at the same time or on the same object). Observations contain values that you measure on different variables.
- A variable is continuous if you can arrange its values in order and an infinite number of values can exist between any two values of the variable.
ggplot2 will treat your variable as continuous if it is a numeric, integer, or a recognizable date-time class (but not a factor
- A variable is discrete if it is not continuous. Discrete variables can only contain a finite (or countably infinite) set of unique values.
ggplot2 will treat your variable as discrete if it is not a numeric, integer, or recognizable date-time class.

##### 5.0.1.0.1 Discrete distributions
```{r}
library(ggplot2)
ggplot(data = diamonds) +
        geom_bar(mapping = aes(x = cut,fill = cut))
#You can compute counts of discrete variables using R's table() function.
table(diamonds$cut)
```

##### 5.0.1.0.2 Continous distributions
> If your data is truly continuous, then no two observations will have the same value—so long as you measure the data precisely enough (e.g. without rounding to the _n_th decimal place).
> To get around this, data scientists divide the range of a continuous variable into equally spaced intervals, a process called binning.
        * They then count how many observations fall into each bin.
        * And display the count as a bar, or some other object.
This method is temperamental because the appearance of the distribution can change dramatically if the bin size changes. As no bin size is “correct,” you should explore several bin sizes when examining data.

Several geoms exist to help you visualize continuous distributions. They almost all use the “bin” stat to implement the above strategy. For each of these geoms, you can set the following arguments for “bin” to use:

        * binwidth - the width to use for the bins in the same units as the x variable
        * origin - origin of the first bin interval
        * right - if TRUE bins will be right closed (e.g. points that fall on the border of two bins will be counted with the bin to the left)
        * breaks - a vector of actual bin breaks to use. If you set the breaks argument, it will override the binwidth and origin arguments.
Use geom_histogram() to make a traditional histogram. The height of each bar reveals how many observations fall within the width of the bar.
```{r}
ggplot(data = diamonds) +
        geom_histogram(aes(x = carat))
# Compare plots
ggplot(data = diamonds) +
        geom_histogram(aes(x = carat), binwidth = 1)

ggplot(data = diamonds) +
  geom_histogram(aes(x = carat), binwidth = 0.01)

# geom_density: displays  densitydensity —not  countcount —on the y axis, which makes it easier to compare the shape of the distributions of multiple subgroups; the area under each curve will be normalized to one, no matter how many total observations occur in the subgroup

ggplot(data = diamonds) +
  geom_density(aes(x = carat, color = cut), kernel = "gaussian", adjust = 4)
```


#### 5.0.2 Compare Distributions
###### 5.0.3 Visualize covariation
5.0.3.1 Visualise functions between two variables
Covariation is a relationship between the values of two or more variables.
However “correlation (covariation) does not prove causation”. 
```{r}
ggplot(data = economics) +
  geom_point(aes(x = date, y = unemploy))

#geom_line makes the relationship clear
ggplot(data = economics) +
  geom_line(aes(x = date, y = unemploy))

#Use geom_step() to turn a line chart into a step function
ggplot(data = economics[1:150, ]) +
  geom_step(aes(x = date, y = unemploy))

# geom_area() creates a line chart with a filled area under the line.
ggplot(data = economics) +
  geom_area(aes(x = date, y = unemploy))
```

###### 5.0.3.1.1 Visualize correlations between two variables

Many variables do not have a functional relationship. As a result, a single value of one variable can correspond to multiple values of another variable.
Height and weight are two variables that are often related, but do not have a functional relationship.
The easiest way to plot the relationship between two variables is with a scatterplot, i.e. geom_point()
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

# jitter adjustment is so useful that ggplot2 provides the geom_jitter()
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy))

# geom_rug() adds a tickmark along each axis for each value observed in the data. 
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_rug(mapping = aes(x = displ, y = hwy), position = "jitter")
```
*Note*
Data set only contains 53,940 points, but the points overplot each other in a way that we cannot fix with jittering.
For large data, it is more useful to plot summary information that describes the raw data than it is to plot the raw data itself. Several geoms can help you do this.

The simplest way to summarise covariance between two variables is with a model line. The model line displays the trend of the relationship between the variables.

Use 'geom_smooth()' to display a model line between any two variables.

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price)) +
  geom_smooth(mapping = aes(x = carat, y = price))

# geom_smooth() will add a loess line to your data if the data contains less than 1000 points, otherwise it will fit a general additive model to your data with a cubic regression spline, and plot the resulting model line.
# Use the method argument of geom_smooth() to add a specific type of model line to your data.
# By default, geom_smooth() will use the formula y ~ x to model your data. You can modify this formula by setting the formula argument to a different formula.

ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price)) +
  geom_smooth(mapping = aes(x = carat, y = price), 
    method = lm, formula = y ~ poly(x, 4))
#avoid being fooled by also inspecting the raw data or calculating the correlation between your variables, e.g. cor(diamonds$carat, diamonds$price)
```
        - `geom_quantile()`
        - `geom_bin2d()` divides the coordinate plane into a two dimensional grid and then displays the number of observations that fall into each bin in the grid.
        - `geom_hex()` works similarly to geom_bin2d(), but it divides the coordinate plane into hexagon shaped bins.

###### 5.0.3.1.2 Visualize correlations between three variables

There are two ways to add three (or more) variables to a two dimensional plot. You can map additional variables to aesthetics within the plot, or you can use a geom that is designed to visualize three variables.

ggplot2 provides three geoms that are designed to display three variables: `geom_raster(),` `geom_tile() and geom_contour()`. These geoms generalize `geom_bin2d() and geom_density()` to display a third variable instead of a count, or a density.

geom_raster() and geom_tile()














