---
title: "Visualization"
author: "Jovani Ntabgoba"
date: "April 29th 2016"
output: 
  pdf_document: 
    latex_engine: xelatex
---
pkgs <- c(
        "broom", "dplyr", "ggplot2", "jpeg", "jsonlite", 
        "knitr", "Lahman", "microbenchmark", "png", "pryr", "purrr", 
        "rcorpora", "readr", "stringr", "tibble", "tidyr"
)
install.packages(pkgs)
library(ggplot2)


```{r cars}
library(knitr)
library(ggplot2)
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```
ggplot(data = <DATA>) + <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
map the colors of your points to the class variable, to reveal class of each car
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r}
#or map the class to the size aesthetic
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
#or map the class to the alpha aesthetic (transparency of points).
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
#or map class to the shape of the points
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```





```{r}
#You can set aesthetic properties manually, e.g make all points become blue
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

if you set the aesthetic:
-inside of the aes() function, ggplot2 will map the aesthetic to data values and build a legend.
-outside of the aes() function, ggplot2 will directly set the aesthetic to your input.

#### Exercises
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = cty, y = hwy > 26, color = class))
```
Use discrete for map color,size,alpha,shape and continuous to map x,y
use displ > 5, kind of staff also.
?geom_point()

### 3.3 Geoms
A geom,a geometrical object that a plot uses to represent data
eg bar charts use bar geoms, boxplot - boxplot geoms, line plot - line geom.
Scatter plots break the flow and use geom_point
```{r}
ggplot(data = mpg) +
        geom_smooth(mapping = aes(x = displ, y = hwy))
#every geom function takes a mapping, 
# not every aesthetic works with every geom
```

```{r}
ggplot(data = mpg) + 
        geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
or to visualize it better, use two geoms
```{r}
ggplot(data = mpg) +
        geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
        geom_smooth(mapping = aes(x = displ, y = hwy, color=drv, linetype = drv))
```






**Note**: ggplot2 provides 37 geom functions. Each suited for a particular type of data or relationship

### 3.4 Layers
To display multiple geoms in the same plot, Add multiple geom functions.
```{r}
ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy)) +
        geom_smooth(mapping= aes(x = displ, y = hwy))
```
to avoid redundancy of code, put the arg in ggplot() to make them global
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
        geom_point() + 
        geom_smooth()
```

NB: Mappings placed instead geom functions are considered local. Can overide global.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
        geom_point(mapping = aes(color = class)) + 
        geom_smooth()
```

You can use the same to specify individual datasets for each layer
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
        geom_point() +
        geom_smooth(data = subset(mpg, cyl == 8))
```

#Exerices
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) + 
        geom_point() + 
        geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
        geom_point() + 
        geom_smooth()
```

up and down plots, are same

```{r}
ggplot(mapping = aes(x = displ, y = hwy)) + 
        geom_point(data = mpg) + 
        geom_smooth(data = mpg)
```

###  3.5 Position adjustments
To make a bar chart, use geom_bar(), doen't require y aesthetic
```{r}
ggplot(data = diamonds) +
        geom_bar(mapping = aes(x = cut))
```

Color aes does outline of bars, fill aes does fill bars
```{r}
ggplot(data = diamonds) +
        geom_bar(mapping = aes(x = cut, fill = cut))
```
### 3.5.1 Position = "identity"
position = "stack", places overappling objects on top of each other, = "fill"
fills the entire bars, letting you easily compare.
position = "dodge", places overlapping objects directly beside one another.
```{r}
ggplot(data = diamonds) +
        geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge") +
        ggtitle('Position ="dodge"')
```
position = "jitter", used for scatter plots, to see weight of points,incase some overlapped
jitter adds a litle bit of noise on each point to make them diffrent if overlapped
```{r}
ggplot(data = mpg) +
        geom_point(mapping = aes(x = displ, y = hwy), position = "jitter") +
        ggtitle('Position = "jitter"')
```

